---
title: 'Regular Expressions based programs'
author: "Yi Chua"
date: "4/17/2020"
output: html_document
---
A small project creating software programs found in everyday life using Regular Expression to match character combinations in strings!


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

```{r}
library(stringr)
```

### Counting Vowels
```{r}
#' @title Counting Vowels
#' @description computes the number of vowels in a character string
#' @param string (character)
#' @return a vector count of the number of vowels
count_vowels <- function(string){
  extract <- str_extract_all(str_to_lower(string), regex('[aeiou]'))
  list_vowels <- c('a','e','i','o','u')
  count <- rep(0,length(list_vowels))
  for (v in (1:length(list_vowels))){
    cum_sum <- cumsum(extract[[1]] == list_vowels[v])
    count[v] <- cum_sum[length(cum_sum)]
  }
  names(count) <- list_vowels
  return(count)
}

#test cases
count_vowels("FIAT LUX")
count_vowels("MaY tHe FoRce Be wIth yOu")
count_vowels("a fox jumps over the lazy dog")
```

### Counting Consonants
```{r}
#' @title Counting Consonants
#' @description computes the number of consonants in a character string
#' @param string (character)
#' @return table count of the number of consonants (vector)
count_consonants <- function(string){
  extract <- str_extract_all(str_to_lower(string), regex('[^aeiou]'))
  list_consonants <- c('b','c','d','f','g','h','j','k','l','m','n','p',
                       'q','r','s','t','v','w','x','y','z')
  count <- rep(0, length(list_consonants))
  for (c in (1:length(list_consonants))){
    cum_sum <- cumsum(extract[[1]] == list_consonants[c])
    count[c] <- cum_sum[length(cum_sum)]
  }
  names(count) <- list_consonants
  return(count)
}

#test cases
count_consonants("FIAT LUX")
count_consonants("MaY tHe FoRce Be wIth yOu")
count_consonants("a fox jumps over the lazy dog")
```



### Hex verifier
```{r}
#' @title Hexadecimal checker
#' @description checks whethter the string is a valid color in hexadecimal notation
#' @param string (character)
#' @return boolean value (Logic)
is_hex <- function(string){
  return(str_detect(str_to_upper(string), regex('^#[A-F0-9]{6}$')))
}

#test case
is_hex("#ff0000")
is_hex("#123456")
is_hex("#12Fb56")
is_hex("#1234GF")
is_hex("#1234567")
is_hex("blue")
```

```{r}
#' @title Hexadecimal w/ Alpha Transparency checker
#' @description checks whethter the string is a valid hex color with alpha transparency
#' @param string (character)
#' @return boolean value (Logic)
is_hex_alpha <- function(string){
  return(str_detect(str_to_upper(string), regex('^#[A-F0-9]{8}$')))
}

#test cases
is_hex_alpha("#FF000078")
is_hex_alpha("#FF0000")
```

### Hexadecimal-RGB Converter
```{r}
#' @title Hexadecimal to RGB Values
#' @description takes a hex-color and returns a named vector with the values of the RGB as well as the alpha channels
#' @param string (character)
#' @return RGB values (vector)
hex_values <- function(string){
  list_rgba <- c("red", "green", "blue", "alpha") #will be used to label the vector with the values of RGB(A)
  rgb <- rep("0",3)
  if (!is_hex(string) & !is_hex_alpha(string)){
  stop("\ninput is not a valid hexadecimal color")
  } else{
    # set pattern depending if it is hex with alpha or without alpha
    if(is_hex_alpha(string)){
    pattern <- '^#([A-F0-9]{2})([A-Fa-f0-9]{2})([A-F0-9]{2})([A-F0-9]{2})$' #pattern for hex w/ alpha
    } else{
      pattern <- '^#([A-F0-9]{2})([A-Fa-f0-9]{2})([A-F0-9]{2})$' #pattern for just hex
    }
    match_group <- str_match(string,regex(pattern)) #pattern matched
    for (m in 2:(length(match_group))){
      rgb[m-1] <-  match_group[m]
      names(rgb)[m-1] <- list_rgba[m-1]
    }
  } 
  return(rgb)
}

#test cases
hex_values("#12345655")
hex_values("#12Fb56")
hex_values("#1234GF")
hex_values("#1234567")
hex_values("blue")
```




### Password verifier
Checks if the string is a "valid" password

We consider a string to be a valid password if it: 

- Has a minimum of 8 characters  
- Contains at least 1 lowercase letter  
- Contains at least 1 uppercase letter  
- Contains at least 1 number   
- Contains at least 1 of the permitted special characters -!@#$%ˆ&*  
- Should not contain pattern *hello* or *123*
```{r}
#' @title Password Checker
#' @description checks if a string is a 'valid' password
#' @param string (character)
#' @return boolean value (logic)
check_pswd <- function(passwords){
return(!(str_detect(passwords, regex("hello"))) & !str_detect(passwords,regex("123")) &str_detect(passwords,regex("^(?=.*[a-z])(?=.*[A-Z])(?=.*[0-9])(?=.*[-!@#$%ˆ&*])(?=.{8,})")))
}


passwords <- c("Hello12!", "Hell@125!", "Hel7o123", "Hel7o123!", "L@R!6373", "Fanny*1", "L@ttie#@*!", "Bern_Zurich")

check_pswd(passwords)
```

### Currency converter
```{r}
#' @title Currency converter
#' @description converts from one currency to another 
#' @param amount numeric value (double)
#' @param from type of currency: euro, pound, yen, yuan, rupee, peso, bitcoin (character)
#' @param to type of currency: euro, pound, yen, yuan, rupee, peso, bitcoin (character)
#' @return the amount based on the type of currency one desired (double)
exchange <- function(amount = 10, from = 'us', to = 'euro'){
  conversion_table <- c("us" = 1,"euro" = 0.87, "pound" = 0.68, "yen" = 106.43, "yuan" = 6.47, "rupee" = 66.38, "peso" = 17.21, "bitcoin" = 0.0022)
  conversion_table
  if ((!(str_extract(from, regex("\\w+"))) %in% names(conversion_table)) | !(str_extract(to, regex("\\w+")) %in% names(conversion_table)))
    stop("currency type must be the following:\nus, euro, pound, yen, yuan, rupee, peso, or bitcoin")
  if (amount < 0){
    stop("\n amount must be nonnegative")
  } else {
    #extract the numerical values that will be used as the currency exchange rate 
    f = conversion_table[which(names(conversion_table) == (str_extract(from, regex("\\w+"))))]
    t = conversion_table[which(names(conversion_table) == (str_extract(to, regex("\\w+"))))]
    converted <- (t / f) * amount
    names(converted) <- NULL
    return(converted)
  }
}

#test cases
exchange(amount = 1, from = 'us', to = 'euro')
exchange(amount = 10, from = 'us', to = 'euro')
exchange(amount = 10, from = 'us', to = 'bitcoin')
exchange(amount = 10, from = 'us', to = 'pound')

#negative amount
exchange(amount = -50, from = 'us', to = 'euro')

#incorrect currency type/nonexistent
exchange(amount = 10, from = 'us', to = 'ringgit')
```

### Dataframe Cleaner
```{r}
#load messy data csv file
dat <- read.csv('USA_crime.csv', header = FALSE, stringsAsFactors = FALSE)
dat
```

```{r}
#' @title Data cleaner
#' @description cleans messy data and return a clean data frame containing information 
#' @param df takes in a data frame (data frame)
#' @return a cleaned up data frame (data frame)
clean_data <- function(df){
  pattern <- "^[\\s]?([A-z]+)[+?>]*[- ]*[\\s]?[>+&$^?:;*=]*(\\d{1,2}\\.\\d{1})[>+&$^?:;*=]*[\\s]*(\\d{3})[>+&$^?:;*=]*[\\s]*(\\d{2})$"
  matched <- str_match(df[[1]],regex(pattern))
  #create empty matrix
  clean_df <- matrix(data= NA, nrow =nrow(matched), ncol = ncol(matched) - 1)
  for (i in 1:nrow(matched)){
    row_df <- matched[i,]
    for (j in 2:ncol(matched)){
      clean_df[i,j-1] <- row_df[j]
    }
  }
  #convert matrix to data frame
  clean_df <- as.data.frame(clean_df,header = FALSE, stringsAsFactors = FALSE)
  names(clean_df) <- c("State", "Murder", "Assault", "UrbanPop")
  return(clean_df)
}

clean_data(dat)
```
